---
output: html_document
jupyter:
  kernelspec:
    name: ir
    display_name: R
    language: R
---

```{r, include = FALSE}
# Show chunks with a `fold` option as collapsible sections with that title
knitr::knit_hooks$set(fold = function(before, options) {
    if (before) {
        return(paste0("\n\n<details><summary>", options$fold, "</summary>\n\n"))
    } else {
        return("\n\n</details>\n\n")
    }
})
```

*(Group members: Mariia Steeghs-Turchina, Henning Lang)*

# Detection of Offensive Tweets

<!-- Team, goal of the competition -->

```{r setup, message = FALSE, code = readLines(here::here("setup.R"))}
source(here::here("setup.R"))
```

```{r notebooks, child = list.files(here::here("analysis"), full.names = TRUE)}
```

# References

Raoul, Karel Veldkamp, and Joost van Kordelaar.
BDA 2026 Detection of Offensive Tweets.
https://www.kaggle.com/competitions/bda-2026-detection-of-offensive-tweets, 2026.
Kaggle.

# Division of work

| Member | Contribution |
|---|---|
| Henning Lang | Feature engineering (e.g. sentiment analysis-based features, offensive or question words), analysis of effect of stopwords exclusion, implementation of the tree-based-xgboost-model (xgboost_tree) , presentation (e.g. objectives of project, questions at the beginning) |
| Mariia Steeghs-Turchina | Feature engineering (e.g. token level features and values, some of the extra features), study of features' effect on performance, implementation of the glmnet models (lasso, ridge, elastic_net), presentation (e.g. structure, data description, model comparison, submission) |
