```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
```

# Get data

## Load raw test and training data

Some observations about the datasets:

- The training dataset consists of 400,000 tweets and contains the response variable label, while the test dataset consists of 50,001 tweets and doesn't contain the response variable.
- The training data contains 3 columns: id, tweet, and label. The test data contains 2 columns: id and tweet.
- Each tweet (character variable) has a unique ID (saved in the variable id).
- The outcome label contains two levels: 1 = offensive language and 0 = no offensive language.
- In the training data approximately half of the tweets contains offensive language.
- Both test and training data contain no missing values.

```{r}
# Find the right file paths
train_filepath <- dir(data_dir, pattern = "train.csv", recursive = TRUE, full.names = TRUE)
test_filepath <- dir(data_dir, pattern = "test.csv", recursive = TRUE, full.names = TRUE)

# Read in the csv files
train_data_raw <- read_csv(train_filepath, col_types = "cci")
test_data_raw <- read_csv(test_filepath, col_types = "cc")

cat("Train data dimensions: ", dim(train_data_raw), "\n")
head(train_data_raw)
cat("Test data dimensions: ", dim(test_data_raw), "\n")
head(test_data_raw)
```

```{r}
train_data <- if (final_submission) train_data_raw else slice_head(train_data_raw, n = 15000)
test_data <- test_data_raw
cat("Using ", nrow(train_data), " training rows and ", nrow(test_data), " test rows.\n")
```

## Cleaning

Remove duplicate tweets, keeping only the first instance of each tweet.

```{r}
train_data <- train_data |>
    distinct(tweet, .keep_all = TRUE)
```

## Split training / validation

We split the tweets before computing features to keep the model unbiased towards validation.

```{r, fold = "Relevant functions for preparing data for models", code = readLines(here::here("R", "data.R"))}
source(here::here("R", "data.R"))
```

```{r}
p <- 0.8
cat("Splitting training data into ", p * 100, "% training and ", (1 - p) * 100, "% validation.\n")
data_split <- split_train_validation(train_data, target_col = "label", partition = p)
```

```{r, include = FALSE}
saveRDS(data_split, here::here(output_dir, "data_split.rds"))
saveRDS(test_data, here::here(output_dir, "test_data.rds"))
```
