```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
data_split <- readRDS(here::here(output_dir, "data_split.rds"))
test_data <- readRDS(here::here(output_dir, "test_data.rds"))
```

# Features

```{r, fold = "Relevant functions for tokenizing", code = readLines(here::here("R", "tokens.R"))}
source(here::here("R", "tokens.R"))
```

```{r, fold = "Relevant functions for feature engineering", code = readLines(here::here("R", "features.R"))}
source(here::here("R", "features.R"))
```

## Potential features

**Notation:**

- $d$: document
- $t$: token
- $n_{d,t}$: count of token $t$ in document $d$
- $N$: total number of documents
- $DF_t$: number of documents containing token $t$

**Token values** (one column per token; `weighting` in `feature_config()`):

| Feature | Equation / representation | Motivation / intuition | Source |
|---|---|---|---|
| **Document occurrence** | $\mathbf{1}(n_{d,t}>0)$ | 0–1 encoding of the presence or absence of a token in a document (here: tweet) | Competition |
| **Token count** | $n_{d,t}$ | Simple counts of each token within documents (resulting in a document by term matrix, or DTM) | Competition |
| **Log count** | $1+\log n_{d,t}$ | Repeats add less than the first occurrence | [Manning et al. (2008)](https://doi.org/10.1017/CBO9780511809071), §6.4.1 |
| **Term frequency (TF)** | $TF_{d,t}=\frac{n_{d,t}}{\sum_{t} n_{d,t}}$ | The relative frequency of a term within a document | Competition |
| **Inverse document frequency (IDF)** | $IDF_t=-\log\left(\frac{DF_t}{N}\right)$ | Inverse of the relative frequency with which a term occurs among the $N$ documents, on a log scale (a measure of 'surprise') | Competition |
| **TF-IDF** | $TFIDF_{d,t}=TF_{d,t}\,IDF_t$ | The product of TF and IDF | Competition; [Spärck Jones (1972)](https://doi.org/10.1108/eb026526) |
| **Naive Bayes ratio** | $\mathbf{1}(n_{d,t}>0)\, r_t$, $r_t=\log\frac{\text{share of offensive tweets with } t}{\text{share of other tweets with } t}$ | Scales each token by how strongly it points to one class | [Wang & Manning (2012)](https://aclanthology.org/P12-2018/); [James et al. (2021)](https://doi.org/10.1007/978-1-0716-1418-1), §4.4.4 |
| **Vector space embeddings** (not used yet) | $\mathbf{z}_d\in\mathbb{R}^k$ | Advanced features like factor loadings (eigenvectors) from a PCA of the DTM, or "word2vec" representations, usually obtained by training neural networks on a very large corpus | Competition |

**Which tokens** (`ngram_max`, `shingle_min`, `shingle_max`, `surprise_pct`):

| Feature | Motivation / intuition | Source |
|---|---|---|
| **Word n-grams** | Keep phrases together: negations, expanded contractions ("is not"), emoji descriptions | [Silge & Robinson (2017)](https://www.tidytextmining.com/ngrams.html), ch. 4; [Davidson et al. (2017)](https://doi.org/10.1609/icwsm.v11i1.14955) |
| **Character n-grams** | Link word forms ("idiot", "idiots") and misspellings | [Waseem & Hovy (2016)](https://doi.org/10.18653/v1/N16-2013); [Nobata et al. (2016)](https://doi.org/10.1145/2872427.2883062) |
| **Surprise filter** | Drop tokens in fewer than `surprise_pct`% of tweets ($IDF_t \le -\log(\text{pct}/100)$) | Competition |

**Extra columns** (`length`, `lexicon`):

| Feature | Motivation / intuition | Source |
|---|---|---|
| **Tweet length** | log number of words, mean word length; weak evidence, may reflect the source dataset | [Nobata et al. (2016)](https://doi.org/10.1145/2872427.2883062); [Waseem & Hovy (2016)](https://doi.org/10.18653/v1/N16-2013) |
| **Sentiment lexicon** | Number of negative / positive words (Bing lexicon); pools rare negative words | [Hu & Liu (2004)](https://doi.org/10.1145/1014052.1014073); [Schmidt & Wiegand (2017)](https://doi.org/10.18653/v1/W17-1101) |

## Options for feature sets

Define options for different feature sets to compare.
Each option is a row in a table, with the columns corresponding to the parameters of `feature_config()` below.

```{r, fold = "Relevant functions for comparing feature sets", code = readLines(here::here("R", "feature_selection.R"))}
source(here::here("R", "feature_selection.R"))
```

### Baseline feature set

```{r}
baseline_key <- "baseline_full_dtm_counts"
baseline <- tribble(
    ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
    baseline_key, 1, 0, 0, "tf", 0, FALSE, FALSE, FALSE
)
```

### Study `weighting`

```{r}
weighting_options <- baseline |>
    bind_rows(
        tribble(
            ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
            "binary", 1, 0, 0, "binary", 0, FALSE, FALSE, FALSE,
            "count", 1, 0, 0, "count", 0, FALSE, FALSE, FALSE,
            "log_count", 1, 0, 0, "log_count", 0, FALSE, FALSE, FALSE,
            "tf", 1, 0, 0, "tf", 0, FALSE, FALSE, FALSE,
            "tf_idf", 1, 0, 0, "tf_idf", 0, FALSE, FALSE, FALSE
        )
    )

study_options <- function(options, title) {
    # display options, evaluate, show and plot result
    cat("Studying", title, "options:\n")
    print(options)
    feature_results <- evaluate_feature_options(options, data_split)
    cat("Results for", title, "options:\n")
    print(feature_results |> select(key, n_features, val_auc, seconds))
    return(feature_results)
}

if (run_studies) {
    study_weighting_results <- study_options(weighting_options, "weighting")
    plot_feature_results(study_weighting_results, baseline = baseline_key)
}
```

On the slice of training data, "binary" weighting is the best option, and it's also the simplest.

```{r}
chosen_weighting <- "binary"
```

### Study `ngram_max`

```{r}
ngram_options <- baseline |>
    bind_rows(
        tribble(
            ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
            "ngrams_2", 2, 0, 0, chosen_weighting, 0, FALSE, FALSE, FALSE,
            "ngrams_3", 3, 0, 0, chosen_weighting, 0, FALSE, FALSE, FALSE,
            "ngrams_5", 5, 0, 0, chosen_weighting, 0, FALSE, FALSE, FALSE,
            "ngrams_10", 10, 0, 0, chosen_weighting, 0, FALSE, FALSE, FALSE
        )
    )

if (run_studies) {
    study_ngram_results <- study_options(ngram_options, "ngram_max")
    plot_feature_results(study_ngram_results, baseline = baseline_key)
}
```

10 ngrams is clearly the best option, but it is also the most expensive in terms of features and computation time.
For the sliced data, it has > 13 times as many features as the next best option (2 ngrams) and takes ~4 times as long to compute.
So, for the sake of speed, we will choose 2 ngrams as the best option for now.

```{r}
chosen_ngram_max <- 2
```

### Study `surprise_pct`

```{r}
surprise_options <- baseline |>
    bind_rows(
        tribble(
            ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
            "surprise_0", chosen_ngram_max, 0, 0, chosen_weighting, 0, FALSE, FALSE, FALSE,
            "surprise_001", chosen_ngram_max, 0, 0, chosen_weighting, 0.01, FALSE, FALSE, FALSE,
            "surprise_005", chosen_ngram_max, 0, 0, chosen_weighting, 0.05, FALSE, FALSE, FALSE,
            "surprise_01", chosen_ngram_max, 0, 0, chosen_weighting, 0.1, FALSE, FALSE, FALSE,
            "surprise_05", chosen_ngram_max, 0, 0, chosen_weighting, 0.5, FALSE, FALSE, FALSE
        )
    )

if (run_studies) {
    study_surprise_results <- study_options(surprise_options, "surprise_pct")
    plot_feature_results(study_surprise_results, baseline = baseline_key)
}
```

The surprise percentage of 0.05 is the best option, although all perform quite similarly until we get to 0.5.
We will stick with 0.05%.

```{r}
chosen_surprise_pct <- 0.05
```

### Study `shingle_min` and `shingle_max`

```{r}
shingle_options <- baseline |>
    bind_rows(
        tribble(
            ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
            "shingles_2", chosen_ngram_max, 2, 2, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_3", chosen_ngram_max, 3, 3, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_4", chosen_ngram_max, 4, 4, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_5", chosen_ngram_max, 5, 5, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_2_5", chosen_ngram_max, 2, 5, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_3_5", chosen_ngram_max, 3, 5, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_4_5", chosen_ngram_max, 4, 5, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_2_4", chosen_ngram_max, 2, 4, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_3_4", chosen_ngram_max, 3, 4, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "shingles_2_3", chosen_ngram_max, 2, 3, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE # nolint: line_length_linter.
        )
    )

if (run_studies) {
    study_shingle_results <- study_options(shingle_options, "shingle_min and shingle_max")
    plot_feature_results(study_shingle_results, baseline = baseline_key)
}
```

The shingles from 2 to 5 characters is the best option, and while it's the most expensive in terms of features and computation time,
the difference is not as extreme, so we will stick with it.

```{r}
chosen_shingle_min <- 2
chosen_shingle_max <- 5
```

### Study `length` and `lexicon`

```{r}
length_lexicon_options <- baseline |>
    bind_rows(
        tribble(
            ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
            "neither", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, FALSE, FALSE, FALSE, # nolint: line_length_linter.
            "length", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, TRUE, FALSE, FALSE, # nolint: line_length_linter.
            "lexicon", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, FALSE, TRUE, FALSE, # nolint: line_length_linter.
            "length_lexicon", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, TRUE, TRUE, FALSE # nolint: line_length_linter.
        )
    )

if (run_studies) {
    study_length_lexicon_results <- study_options(length_lexicon_options, "length and lexicon")
    plot_feature_results(study_length_lexicon_results, baseline = baseline_key)
}
```

Adding the length and/or lexicon columns does not improve the validation AUC,
so we will not include them in the final feature set.

```{r}
chosen_length <- FALSE
chosen_lexicon <- FALSE
```

### Final set of features

```{r}
feature_options <- tribble(
    ~key, ~ngram_max, ~shingle_min, ~shingle_max, ~weighting, ~surprise_pct, ~length, ~lexicon, ~nb_scaling, # nolint: line_length_linter.
    "binary", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, chosen_length, chosen_lexicon, FALSE, # nolint: line_length_linter.
    "nb", chosen_ngram_max, chosen_shingle_min, chosen_shingle_max, chosen_weighting, chosen_surprise_pct, chosen_length, chosen_lexicon, TRUE # nolint: line_length_linter.
)
feature_options
```

## Build the chosen feature sets

Each chosen feature set is built for the training and validation tweets,
so the models can be compared on all of them.

```{r}
feature_sets <- feature_options$key |>
    set_names() |>
    map(\(feature_key) {
        config <- config_from_options(feature_options, feature_key)
        train <- prepare_features(data_split$train, config)
        val <- prepare_features(data_split$val, recipe = train$recipe)

        return(list(config = config, train = train[c("x", "y")], val = val[c("x", "y")]))
    })

tibble(
    feature_set = names(feature_sets),
    n_features = map_int(feature_sets, \(feature_set) ncol(feature_set$train$x))
)
```

```{r, include = FALSE}
saveRDS(feature_sets, here::here(output_dir, "feature_sets.rds"))
```
