```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
feature_sets <- readRDS(here::here(output_dir, "feature_sets.rds"))
```

# Models

The models are compared using held-out validation AUC. Within the training data,
lambda for the glmnet models and the number of boosting rounds (nrounds) for xgboost
are selected using three-fold cross-validation with AUC as the performance measure.

| Method | Model | Package | Tune | Notes |
|---|---|---|---|
| `lasso` | Lasso logistic regression | glmnet | `lambda` | `alpha = 1` |
| `ridge` | Ridge logistic regression | glmnet | `lambda` | `alpha = 0` |
| `elastic_net` | Elastic net logistic regression | glmnet | `lambda` | `alpha = 0.5` |
| `xgboost_tree` | Gradient-boosted decision-tree classifier | xgboost | `nrounds` | fixed tree settings (see xgboost_params) |

> *Note:* The xgboost-package is not compliant with the competition rules,
> since it wasn't mentioned in the ISLR book nor in class (yet).
> In consultation with Raoul we were allowed to use it in our code,
> provided that we do not use the model as the best model for the leaderboard submission.

## Define options for training models

| Column | Meaning |
|---|---|
| `key` | Name of the model |
| `feature_set` | Key of a feature set from the features notebook |
| `method` | `lasso`, `ridge`, `elastic_net`, or `xgboost_tree` |

```{r}
model_options <- tribble(
    ~feature_set, ~method,
    "binary", "lasso",
    "binary", "ridge",
    "binary", "elastic_net",
    "binary", "xgboost_tree",
) |>
    mutate(key = paste(method, feature_set, sep = "_"))

cat("Model options:\n")
model_options
```

## Fit models

Each model is tuned by cross-validation on the training tweets (the same 3 folds for every model),
and scored on the validation tweets.

```{r, fold = "Relevant functions for fitting models", code = readLines(here::here("R", "models.R"))}
source(here::here("R", "models.R"))
```

```{r}
cat("Fitting models...\n")
fitted_models <- fit_all(model_options, feature_sets)
cat("Results of fitted models:\n")
fitted_models$results
```

## Evaluation

The validation AUC of each model.

```{r}
plot_model_auc(fitted_models$results)
```

```{r}
plot_feature_results(
    fitted_models$results,
    baseline_key = "lasso_binary", title = "Validation AUC per model"
)
```

How each model was tuned:

```{r}
plot_tuning(fitted_models$models)
```

```{r}
plot_roc(fitted_models$models, feature_sets)
```

## Find the best model

Choose the best model based on the highest validation AUC and refit it on all labelled tweets (training + validation).

```{r}
best_key <- fitted_models$results |>
    slice_max(val_auc, n = 1, with_ties = FALSE) |>
    pull(key)
cat("Best model:", best_key, "\n")

cat("Refitting the best model on all labelled tweets (training + validation)...\n")
best_fit <- refit_best(best_key, model_options, feature_sets, fitted_models$models)
```

> *Note:* As the xgboost-model wasn't the best model, we don't have to exclude it manually in the following code
> (because we are not allowed to submit this model, as discussed).

## Most important features

We investigate which features influence the prediction of the best model towards most and least offensive.
The coefficients are standardized (multiplied by the standard deviation of their column) to compensate for the different scales of the features.

```{r}
plot_important_words(best_fit$model, feature_sets[[best_fit$model$feature_set]]$train$x, 30, 15)
```

```{r, include = FALSE}
saveRDS(best_fit, here::here(output_dir, "best_fit.rds"))
```
