```{r, include = FALSE}
source(here::here("setup.R"))
for (file in list.files(here::here("R"), full.names = TRUE)) source(file)
feature_sets <- readRDS(here::here(output_dir, "feature_sets.rds"))
```

# Models

Methods:

1. Lasso and Ridge regression
2. Principal Components and Partial Least Squares regression
3. Smoothing
4. Regression and Classification trees (CART)
5. Random Forests
6. Support Vector Machines

Methods (1) and (2) are able to take many features while automatically reducing redundant flexibility to any desired level.
Multicollinearity (highly correlated linear combinations of features) is also automatically taken care of by these methods.

Number (3) on the list, smoothing, grants more flexibility by allowing for some non-linearity in the relations between features and the target variable,
without the need to manually specify a specific mathematical form (as is necessary in polynomial regression).

Methods (4), (5), and (6) are not only able to remove redundant features, but also can automatically recognize interactions between features.

We currently fit:

| Method | Package | Tune | Notes |
|---|---|---|---|
| `lasso` | glmnet | `lambda` | `alpha = 1` |
| `ridge` | glmnet | `lambda` | `alpha = 0` |
| `elastic_net` | glmnet | `lambda` | `alpha = 0.5` |

TODO: SVM and others

## Define options for training models

| Column | Meaning |
|---|---|
| `key` | Name of the model |
| `feature_set` | Key of a feature set from the features notebook |
| `method` | Name of a method in `model_methods` |

glmnet standardizes the columns, which cancels the naive Bayes scaling,
so the glmnet methods are only fitted on the `binary` feature set.

```{r}
model_options <- tribble(
    ~feature_set, ~method,
    "binary", "lasso",
    "binary", "ridge",
    "binary", "elastic_net"
) |>
    mutate(key = paste(method, feature_set, sep = "_"))
model_options
```

## Fit models

Each model is tuned by cross-validation on the training tweets (the same 3 folds for every model),
and scored on the validation tweets.

```{r, fold = "Relevant functions for fitting models", code = readLines(here::here("R", "models.R"))}
source(here::here("R", "models.R"))
```

```{r}
fitted_models <- fit_all(model_options, feature_sets)
fitted_models$results
```

## Evaluation

```{r}
plot_feature_results(
    fitted_models$results,
    baseline_key = "lasso_binary", title = "Validation AUC per model"
)
```

How each model was tuned:

```{r}
plot_tuning(fitted_models$models)
```

```{r}
plot_roc(fitted_models$models, feature_sets)
```

## Find the best model

Choose the best model based on the highest validation AUC and refit it on all labelled tweets (training + validation).

```{r}
best_key <- fitted_models$results |>
    slice_max(val_auc, n = 1, with_ties = FALSE) |>
    pull(key)
cat("Best model:", best_key, "\n")

best_fit <- refit_best(best_key, model_options, feature_sets, fitted_models$models)
```

```{r, include = FALSE}
saveRDS(best_fit, here::here(output_dir, "best_fit.rds"))
```
